Thermodynamic fragments for magic states in quantum computation

### 2021-03-12 magic_paper.tex
Added comparisons with prakash

In particular, any $\sigma$--fragment, with $\sigma$ not full-rank, can be approximated by the $\sigma'$--fragment defined as a mixture of $\sigma$ with uniform noise, $\sigma' = (1-\epsilon)\sigma + \epsilon\frac{1}{d}\id$, for some infinitesimal $\epsilon > 0$, so that $\sigma'$ is full-rank and arbitrarily close to $\sigma$.
Important examples include pure stabilzer states which are rank-$1$ states e.g. the zero state depicted in~\cref{fig:zero}. 
Operations in such fragments include important stabilizer operations like the replacement channel, $\E(\rho) = \sigma$ for all states $\rho$.

@@

In the sections~\cref{sec:unital,sec:stab}, we study majorization constraints on magic distillation arising in different fragments. 
For this reason, we define the $n$-copy, $\epsilon$-noisy Strange state,
\begin{equation}\label{eq:noisysn}
    \rho_{\rm{S}}(\epsilon)^{\otimes n} \coloneqq \left[ (1 - \epsilon) \ket{\rm{S}}\bra{\rm{S}} + \epsilon \frac{1}{3}\id \right]^{\otimes n},
\end{equation}
in $\O_\sigma$, where $\ket{\rm{S}}$ is defined in~\cref{fig:strange} and the induced noise is uniform.

@@

We focus on the distillation process
\begin{equation}\label{eq:stdist}
	\rho_{\rm{S}}(\epsilon)^{\otimes n} \longrightarrow \rho_{\rm{S}}(\epsilon')^{\otimes n'} \otimes \gamma_\beta^{\otimes (n-n')},
\end{equation}
where the noisy Strange state is given in~\cref{eq:noisysn}, but more general states can be handled in exactly the same manner.
We also highlight that any magic state can be written in terms of a noisy Strange state.
Notice that tensoring in copies of $\gamma_\beta$ does not affect the process, since the state is free.
In the following result, all copies $n, n', n - n'$ are even with $n > n'$, and $\epsilon'=0$, but the bounds are easily generalised to odd numbers of copies and $\epsilon'$ such that $0 < \epsilon' < \epsilon$. 
Finally, we assume a \nick{generous} range of initial noise levels, $\epsilon \leq 3/7$, so that the largest Wigner component of the 1-copy noisy magic state is negative.

@@

Consider a qutrit system with maximally Wigner-negative magic state $\ket{{\rm{S}}} = \frac{1}{\sqrt{2}}(\ket{1} - \ket{2})$.
It can be written in density operator form as
\begin{align}
	\ketbra{\rm{S}} = -\frac{1}{3} A_{\bmo} + \sum_{\bmx \neq \bmo} \frac{1}{6} A_{\bmx},
\end{align}
where $\bmx$ ranges in the $3$-dimensional phase space $\cal{P}_d$.
This is the pure Strange state visualised in~\cref{fig:strange}.
We allow for isotropic (uniform) noise, so that the system state is a convex sum of the pure Strange state and the maximally mixed state,
\begin{equation}
	\rho_{\rm{S}}(\epsilon) \coloneqq (1-\epsilon)\ketbra{\rm{S}} + \epsilon \frac{1}{3}\id,
\end{equation}
where we choose to restrict the noise level, $0 \leq \epsilon \leq \frac{3}{7}$, so that the Wigner distribution of the state always contains a negative component which is greater than the remaining 8 positive components.

@@

In all three scenarios, it is simple to check that the first elbow of the initial state is always located closer to $0$ compared to the first elbow of the target state, $x_0 \leq x_0'$:
\begin{enumerate}
	\item $\rm{(C1)} \rightarrow \rm{(C1)}$. 
	It is always true that $e^{-\beta E_0} / \Z_\beta \leq 1$, so
	\begin{align*}
		x_0 &= \left(\dfrac{e^{-\beta E_0}}{3\Z_\beta}\right)^n = \left(\dfrac{e^{-\beta E_0}}{3\Z_\beta}\right)^{n-n'} \left(\dfrac{e^{-\beta E_0}}{3\Z_\beta}\right)^{n'} \\
		&< \left(\dfrac{e^{-\beta E_0}}{3\Z_\beta}\right)^{n'} = x_0'		
	\end{align*}
	\item $\rm{(C2)} \rightarrow \rm{(C1)}$.
	It is always true that
	\begin{align*}
		&\Z_\beta^{n-n'} > (e^{-\beta E_{\rm{max}}})^{n-n'} = e^{-\beta (n-n') E_{\rm{max}}}, \text{ so} \\
		x_0 &= \left(\dfrac{e^{-\beta E_{\rm{max}}}}{\Z_\beta}\right)^n = \dfrac{e^{-\beta(n E_{\rm{max}} - n' E_0)}}{\Z_\beta^{n-n'}} \left(\dfrac{e^{-\beta E_0}}{3\Z_\beta}\right)^{n'} \\
		&< \dfrac{e^{-\beta(n E_{\rm{max}} - n' E_0)}}{e^{-\beta (n-n') E_{\rm{max}}}} \left(\dfrac{e^{-\beta E_0}}{3\Z_\beta}\right)^{n'} \\
		&= e^{-\beta n' (E_{\rm{max}} - E_0)} \left(\dfrac{e^{-\beta E_0}}{3\Z_\beta}\right)^{n'} \\
		&< \left(\dfrac{e^{-\beta E_0}}{3\Z_\beta}\right)^{n'} = x_0'.
	\end{align*}
	\item $\rm{(C2)} \rightarrow \rm{(C2)}$.
	It is always true that $e^{-\beta E_{\rm{max}}} / \Z_\beta \leq 1$, so
	\begin{align*}
		x_0 &= \left(\dfrac{e^{-\beta E_{\rm{max}}}}{\Z_\beta}\right)^n = \left(\dfrac{e^{-\beta E_{\rm{max}}}}{\Z_\beta}\right)^{n-n'} \left(\dfrac{e^{-\beta E_{\rm{max}}}}{\Z_\beta}\right)^{n'} \\
		&\leq \left(\dfrac{e^{-\beta E_{\rm{max}}}}{\Z_\beta}\right)^{n'} = x_0'	
	\end{align*}
\end{enumerate}
We can thus find the target state Lorenz curve coordinate $L'_\star$ at location $x_\star$ by interpolating between the origin and the target state's first elbow, 
\begin{equation}\label{eq:constraint_deriv_temp}
	L'_\star = \frac{x_\star}{x'_0}L'_0.
\end{equation}
We need $L_0 \geq L_\star'$. Therefore, restricting to the first elbow, the necessary distillation condition in~\cref{eq:first_elbow_constraint} simply becomes
\begin{equation}\label{eq:first_elb_bound}
	\frac{L_0}{x_0} \geq \frac{L_0'}{x_0'}.
\end{equation}
Involving more elbows gives stricter, but more convoluted necessary distillation constraints.

@@

@@

An $n$-copy state $\rho^n$ has \emph{error rate} $\epsilon$ iff $\rho$ has an overlap of at least $1-\epsilon$ with magic state $\ket{\rm{S}}$,
\begin{equation}
	\bra{\rm{S}}\rho\ket{\rm{S}} \geq 1-\epsilon.
\end{equation}
The $\epsilon$-noisy Strange state $\rho_{\rm{S}}(\epsilon)^{\otimes n}$ has error rate $\frac{2}{3}\epsilon$.
A general state of $n$ qutrits with error rate $\epsilon$ can be converted to the form $\rho_{\rm{S}}(\epsilon_0)^{\otimes n}$ for some noise level $\epsilon_0$ not necessarily equal to $\epsilon$.

Given an $[[n,n',d]]$ error correcting code, a distillation protocol that achieves the process in~\cref{eq:test} requires
\begin{align}
	\epsilon' &\sim O\left((n\epsilon)^d\right) \\
	\frac{n}{n'} &\sim O(\log^\gamma{(1/\epsilon')}),\ \gamma = \log_d{\left(\frac{n}{n'}\right)}
\end{align}

In~\cite{Bravyi2012,Hastings2018}, the asymptotic rate $O(\log^\gamma{(1/\epsilon')})$ is called the \emph{distillation cost}.

There are many distillation protocols in the literature, e.g.~\cite{Bravyi2012,Hastings2018}, that involve a specific value of $\gamma$.
They should all satisfy
\begin{equation}\label{eq:rate_cost}
	\frac{1}{R(\epsilon, \epsilon', \beta)} \leq \frac{n}{n'} \sim O(\log^\gamma{(1/\epsilon')}).
\end{equation}
Therefore, given process parameters $n, n', \epsilon, \epsilon'$, we can scan through the stabilizer states to find a range of $\beta$ for which inequality~\cref{eq:rate_cost} is valid.
In principle, the higher the value of parameter $\gamma$, the lower the output error rate $\epsilon'$ can be while~\cref{eq:rate_cost} is still true.
In~\cite{Hastings2018} \nick{CITE MORE}, they report a value of $\gamma = 0.6779\dots$, while in~\cite{Krishna2019}, they report a scaling $\gamma = O(1/\log{d})$ with respect to the system's odd prime dimension $d$.
The latter tends to 0 asymptotically.

Furthermore, in~\cite{cit:bravyi,cit:prakash}, two different distillation protocols are proposed with output error rate scalings of the form 
\begin{equation}
	\epsilon' \propto (\alpha \epsilon)^{n^\xi},
\end{equation}
where $\alpha$ is some constant, $\xi = 1/\log_3{15} \approx 0.4$ in~\cite{cit:bravyi} for qubit codes and $\xi = 1/\log_3{19008} \approx 0.1$ for qutrit codes.
They also report a certain input threshold error rate $\epsilon_\star$ above which distillation is not possible.
A first requirement is that $\epsilon_\star \leq E(n'/n, \epsilon', \beta)$, but we can also again find a range of $\beta$ for which the inequality
\begin{equation}
	\epsilon \leq E\left(\frac{n'}{n}, \epsilon'(\epsilon), \beta \right)
\end{equation}
holds, given some distillation rate $n'/n$ and input error rate $\epsilon$.

>>>>

In particular, any $\sigma$--fragment, with $\sigma$ not full-rank, can be approximated by the $\sigma'$--fragment defined as a mixture of $\sigma$ with depolarising noise, $\sigma' = (1-\epsilon)\sigma + \epsilon\frac{1}{d}\id$, for some infinitesimal $\epsilon > 0$, so that $\sigma'$ is full-rank and arbitrarily close to $\sigma$.
Important examples include pure stabilzer states which are rank-$1$ states e.g. the zero state depicted in~\cref{fig:zero}. 
Operations in such fragments include important stabilizer operations like the replacement channel, $\E(\rho) = \sigma$ for all states $\rho$.

@@

In the sections~\cref{sec:unital,sec:stab}, we study majorization constraints on magic distillation arising in different fragments. 
For this reason, we define the $n$-copy, $\epsilon$-noisy Strange state,
\begin{equation}\label{eq:noisysn}
    \rho_{\rm{S}}(\epsilon)^{\otimes n} \coloneqq \left[ (1 - \epsilon) \ket{\rm{S}}\bra{\rm{S}} + \epsilon \frac{1}{3}\id \right]^{\otimes n},
\end{equation}
in $\O_\sigma$, where $\ket{\rm{S}}$ is coined in~\cref{fig:strange} and we induce uniform depolarising noise.

@@

We focus on the distillation process
\begin{equation}\label{eq:stdist}
	\rho_{\rm{S}}(\epsilon)^{\otimes n} \longrightarrow \rho_{\rm{S}}(\epsilon')^{\otimes n'} \otimes \gamma_\beta^{\otimes (n-n')},
\end{equation}
where the noisy Strange state is given in~\cref{eq:noisysn}, but more general states can be handled in exactly the same manner.
We also highlight that any magic state can be written in terms of a noisy Strange state.
Notice that tensoring in copies of $\gamma_\beta$ does not affect the process, since the state is free.
In the following result, all copies $n, n', n - n'$ are even with $n > n'$, and $\epsilon'=0$, but the bounds are easily generalised to odd numbers of copies and $\epsilon'$ such that $0 < \epsilon' < \epsilon$. 
Finally, we assume a range of initial noise levels, $\epsilon \leq 3/7$  \nick{higher than all threshold values CITE}, so that the largest Wigner component of the 1-copy noisy magic state is negative.

@@

Consider a qutrit system with maximally Wigner-negative magic state $\ket{{\rm{S}}} = \frac{1}{\sqrt{2}}(\ket{1} - \ket{2})$.
It can be written in density operator form as
\begin{align}
	\ketbra{\rm{S}} = -\frac{1}{3} A_{\bmo} + \sum_{\bmx \neq \bmo} \frac{1}{6} A_{\bmx},
\end{align}
where $\bmx$ ranges in the $3$-dimensional phase space $\cal{P}_d$.
This is the pure Strange state visualised in~\cref{fig:strange}.
We allow for depolarising noise, so that the system state is a convex sum of the pure Strange state and the maximally mixed state,
\begin{equation}
	\rho_{\rm{S}}(\epsilon) \coloneqq (1-\epsilon)\ketbra{\rm{S}} + \epsilon \frac{1}{3}\id,
\end{equation}
where we choose to restrict the noise level, $0 \leq \epsilon \leq \frac{3}{7}$, so that the Wigner distribution of the state always contains a negative component which is greater than the remaining 8 positive components.

@@

In all three scenarios, it is simple to check that the first elbow of the initial state is always located closer to $0$ compared to the first elbow of the target state, $x_0 \leq x_0'$:
\begin{enumerate}
	\item $\rm{(C1)} \rightarrow \rm{(C1)}$. 
	It is always true that $e^{-\beta E_0} / \Z_\beta \leq 1$, so
	\begin{align*}
		x_0 &= \left(\dfrac{e^{-\beta E_0}}{3\Z_\beta}\right)^n = \left(\dfrac{e^{-\beta E_0}}{3\Z_\beta}\right)^{n-n'} \left(\dfrac{e^{-\beta E_0}}{3\Z_\beta}\right)^{n'} \\
		&< \left(\dfrac{e^{-\beta E_0}}{3\Z_\beta}\right)^{n'} = x_0'		
	\end{align*}
	\item $\rm{(C2)} \rightarrow \rm{(C1)}$.
	It is always true that
	\begin{align*}
		&\Z_\beta^{n-n'} > (e^{-\beta E_{\rm{max}}})^{n-n'} = e^{-\beta (n-n') E_{\rm{max}}}, \text{ so} \\
		x_0 &= \left(\dfrac{e^{-\beta E_{\rm{max}}}}{\Z_\beta}\right)^n = \dfrac{e^{-\beta(n E_{\rm{max}} - n' E_0)}}{\Z_\beta^{n-n'}} \left(\dfrac{e^{-\beta E_0}}{3\Z_\beta}\right)^{n'} \\
		&< \dfrac{e^{-\beta(n E_{\rm{max}} - n' E_0)}}{e^{-\beta (n-n') E_{\rm{max}}}} \left(\dfrac{e^{-\beta E_0}}{3\Z_\beta}\right)^{n'} \\
		&= e^{-\beta n' (E_{\rm{max}} - E_0)} \left(\dfrac{e^{-\beta E_0}}{3\Z_\beta}\right)^{n'} \\
		&\leq \left(\dfrac{e^{-\beta E_0}}{3\Z_\beta}\right)^{n'} = x_0'.
	\end{align*}
	\item $\rm{(C2)} \rightarrow \rm{(C2)}$.
	It is always true that $e^{-\beta E_{\rm{max}}} / \Z_\beta \leq 1$, so
	\begin{align*}
		x_0 &= \left(\dfrac{e^{-\beta E_{\rm{max}}}}{\Z_\beta}\right)^n = \left(\dfrac{e^{-\beta E_{\rm{max}}}}{\Z_\beta}\right)^{n-n'} \left(\dfrac{e^{-\beta E_{\rm{max}}}}{\Z_\beta}\right)^{n'} \\
		&\leq \left(\dfrac{e^{-\beta E_{\rm{max}}}}{\Z_\beta}\right)^{n'} = x_0'	
	\end{align*}
\end{enumerate}
We can thus find the target state Lorenz curve coordinate $L'_\star$ at location $x_\star$ by interpolating between the origin and the target state's first elbow, 
\begin{equation}\label{eq:constraint_deriv_temp}
	L'_\star = \frac{x_\star}{x'_0}L'_0.
\end{equation}
We need $L_0 \geq L_\star'$. Therefore, restricting to the first elbow, the necessary distillation condition in~\cref{eq:first_elbow_constraint} simply becomes
\begin{equation}\label{eq:first_elb_bound}
	\frac{L_0}{x_0} \geq \frac{L_0'}{x_0'}.
\end{equation}
Involving more elbows gives stricter, but more convoluted necessary distillation constraints.

@@

We apply our results on the ternary Golay code distillation protocol suggested in~\cite{cit:prakash}.
In this protocol, a single Strange state is distilled with noise level
\begin{equation}
	\epsilon'(n, \epsilon) \approx \frac{1}{\alpha}(\alpha\epsilon)^{n^\xi},
\end{equation}
where $\alpha \approx 1.75$ and $\xi \approx 0.112$ are constants specific to the protocol.
This can be rewritten in terms of the protocol distillation rate,
\begin{equation}
	R_{\rm{Golay}}(\epsilon, \epsilon') = \frac{1}{n} \approx \left( \frac{\log{(\alpha \epsilon)}}{\log{(\alpha \epsilon')}} \right)^{\frac{1}{\xi}}
\end{equation}

We compare the distillation rate of the protocol with our distillation bound, 
\begin{equation}
	R_{\rm{Golay}}(\epsilon, \epsilon') \leq R(\epsilon, \epsilon', \beta)
\end{equation} 
in figure~\cref{fig:prakash_compare}
\begin{figure}[h]
    \centering
    \includegraphics[scale=0.5]{figs/prakash_compare.png}
    \caption{\textbf{Comparison with ternary Golay distillation rate.}
    Our bound is not violated but gets close to saturated for some noise levels.
    The thick diagonal line indicates where $\epsilon' = \epsilon$.
    }
    \label{fig:prakash_compare}
\end{figure}

We can also rewrite the expression as 
\begin{equation}
	\epsilon_{\rm{Golay}}(\epsilon', n) \approx \frac{1}{\alpha}(\alpha\epsilon')^{n^{-\xi}},
\end{equation}
and compare with the numerically optimal error bounds provided by majorization as shown in~\cref{fig:prakash_compare_errors}
\begin{figure}[h]
    \centering
    \includegraphics[scale=0.5]{figs/prakash_compare_errors.png}
    \caption{\textbf{Comparison with ternary Golay output error rate.}
    Our bound is not violated but achieves a maximum saturation of $\epsilon_\star - \epsilon_{\rm{Golay}} = 0.147$.
    }
    \label{fig:prakash_compare_errors}
\end{figure}

\iffalse
An $n$-copy state $\rho^n$ has \emph{error rate} $\delta$ iff $\rho$ has an overlap of at least $1-\delta$ with magic state $\ket{\rm{S}}$,
\begin{equation}
	\bra{\rm{S}}\rho\ket{\rm{S}} \geq 1-\delta.
\end{equation}
The $\epsilon$-noisy Strange state $\rho_{\rm{S}}(\epsilon)^{\otimes n}$ has error rate $\delta = \frac{2}{3}\epsilon$.
A general state of $n$ qutrits with error rate $\delta$ can be converted to the form $\rho_{\rm{S}}(\epsilon_0)^{\otimes n}$ for some noise level $\epsilon$ not necessarily equal to $\delta$.

Given an $[[m,k,d]]$ error correcting code, a distillation protocol that achieves the process in~\cref{eq:test} requires that the following bound holds,
\begin{align}
	\frac{n}{n'} = O(\log^\gamma{(1/\epsilon')}),\ \gamma = \log_d{\left(\frac{m}{k}\right)}
\end{align}

In~\cite{Bravyi2012,Hastings2018}, the asymptotic rate $O(\log^\gamma{(1/\epsilon')})$ is called the \emph{distillation cost}, and in~\cite{cit:bravyi,Krishna2019} the parameter $\gamma$ is called the \emph{overhead}.

There are many distillation protocols in the literature, e.g.~\cite{Bravyi2012,Hastings2018}, that involve a specific value of $\gamma$ and convert a specific number of copies $n$ of input states to $n'=1$ output state \nick{what about $\epsilon$?}.
They should all satisfy
\begin{equation}\label{eq:rate_cost}
	\frac{1}{R(\epsilon, \epsilon', \beta)} \leq \frac{n}{n'} = O(\log^\gamma{(1/\epsilon')}).
\end{equation}
Therefore, given process parameters $n, n', \epsilon, \epsilon'$, we can scan through the stabilizer states to find a range of $\beta$ for which inequality~\cref{eq:rate_cost} is valid.
In principle, the higher the value of parameter $\gamma$, the lower the output error rate $\epsilon'$ can be while~\cref{eq:rate_cost} is still true.
In~\cite{Hastings2018} \nick{CITE MORE}, they report a value of $\gamma = 0.6779\dots$, while in~\cite{Krishna2019}, they report a scaling $\gamma = O(1/\log{d})$ with respect to the system's odd prime dimension $d$.
The latter tends to 0 asymptotically.

@@

Furthermore, in~\cite{cit:bravyi,cit:prakash}, two different distillation protocols are proposed with output error rate scalings of the form 
\begin{equation}
	\epsilon' \propto (\alpha \epsilon)^{n^\xi},
\end{equation}
where $\alpha$ is some constant, $\xi = 1/\log_3{15} \approx 0.4$ in~\cite{cit:bravyi} for qubit codes and $\xi = 1/\log_3{19008} \approx 0.1$ for qutrit codes.
They also report a certain input threshold error rate $\epsilon_\star$ above which distillation is not possible.
A first requirement is that $\epsilon_\star \leq E(n'/n, \epsilon', \beta)$, but we can also again find a range of $\beta$ for which the inequality
\begin{align}
	\epsilon &\leq E\left(\frac{n'}{n}, \epsilon'(\epsilon), \beta \right) \\
	\epsilon' &\sim O\left((n\epsilon)^d\right)
\end{align}
holds, given some distillation rate $n'/n$ and input error rate $\epsilon$.
\fi
